### 2011-07-31 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

\chapter{Analysis of Recent Approaches}
\todo[inline]{Translate.}
Existuje mnoho prístupov generovania XML schém. Väčšina z nich spracováva XML data \todo{spomenúť alebo rozpísať tieto prístupy, ako napríklad EXTRACT?}. Podľa mojej znalosti v dobe písania práce existuje len jediný \todo{je pravda?} prístup využívajúci XML operácie\todo{ref do literatury}. Ten sa snaží hľadať keys a foreign keys za pomoci XQuery dotazov.

\section{Discovering XML Keys and Foreign Keys in Queries}
Táto metóda sa snaží vylepšiť automatické odvodzovanie XML schém hľadaním keys a foreign keys zo sady XQuery dotazov. Pre toto hľadanie sú použité len tieto dotazy, nie samotné XML data. Výsledkom je zoznam keys a foreign keys, ktoré môžu byť zachytené pomocou \textbf{key}, \textbf{keyref} \todo{aj unique?} konštrukcií jazyka XML Schema.

\subsection{Keys and foreign keys}
Autori si zavádzajú definíciu pre key a foreign key:

\begin{define}[Key]
Key je trojica $$(C, P, \{L\})$$ kde $C$, $P$, $L$ sú XPath cesty bez predikátov a používajú len osi \texttt{child} a \texttt{descendant}.
$C$ môže byť vynechané. $(P, \{L\})$ je v takom prípade to isté ako $(/, P, \{L\})$ a nazýva sa \emph{global key}, inak sa nazýva \emph{local key}.
\end{define}

@@

\begin{define}[Foreign key]
Foreign key je konštrukcia $$(C, (P_1, \{L_1\}) \Rightarrow (P_2, \{L_2\}))$$ kde $(C, P_2, \{L_2\})$ je key, $P_1$, $L_1$ sú XPath cesty bez predikátov a používajú len osi \texttt{child} a \texttt{descendant}.
$C$ môže byť vynechané podobne ako v prípade key.
\end{define}

@@

\subsection{Key and foreign key discovery}
\subsubsection{Joins in queries}
Na hľadanie využíva metóda element/element joins v dotazoch.

@@

Metóda predpokladá, že každý join je robený cez pár key/foreign key. On the base of this assumption, it can be inferred from $Q$ that $L_1$ is a key for elements in $S_1$ or $L_2$ is a key for elements in $S_2$ and the other is a foreign key referencing the key.

Autori uvádzajú dve pozorovania:
\begin{enumerate}
\renewcommand{\theenumi}{(O\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\item Assume that an element $e_1$ from $S_1$ can be joined with more elements from $S_2$. It means that there can be more different elements in $S_2$ having their value of $L_2$ equal to $e_1/L_1$. In other words, there can be more different elements in $S_2$ having the same value of $L_2$. Therefore, we infer that $L_2$ can not be a key of the elements in $S_2$. Moreover, because we suppose that one of $L_1$ and $L_2$ is a key and $L_2$ can not be a key, we infer that $L_1$ is a key of the elements in $S_1$. Consequently, we infer that $L_2$ is a foreign key referring $L_1$.
\item Vice versa, assume that $e_1$ can be joined with maximally one element from $S_2$. It means that there is maximally one element in $S_2$ having its value of $L_2$ equal to $e_1/L_1$. In that case we suppose that each element in $S_2$ has a unique value of $L_2$. Therefore, we infer that $L_2$ is a key of the elements in $S_2$ and consequently that $L_1$ is a foreign key referring $L_2$. We cannot infer whether $L_1$ is a key or not.
\end{enumerate}

\subsubsection{Join patterns}
Rozhodnutie pre jeden z prípadov (O1) a (O2) pre konrétny join je robené podľa tvaru tohto joinu. V dotaze sa hľadajú takzvané \emph{join patters}. Tie sú dva, \texttt{for} \emph{join pattern} a \texttt{let} \emph{join pattern} a vyzerajú takto (prvý je \texttt{for} \emph{join pattern}, druhý \texttt{let} \emph{join pattern}):
\todo[inline]{spravne kreslit indexy pri pismenkach}
\begin{verbatim}
01  for e1 in P1               for $e1 in P1
02  return                     return
03    for $e2 in                 let $e2 :=
        P2[L2 = $e1/L1]            P2[L2 = $e1/L1]
04    return CR                  return CR
\end{verbatim}

@@

\section{Scoring function}
Môže nastať situácia, že nie je splnený nejaký z predpokladov, z ktorých metóda vychádza. Predpokladajme nejakú množinu dotazov. Môže sa stať, že z nejakého dotazu z tejto množiny metóda odvodí key $K$ a z nejakého iného dotazu, že $K$ nie je key. \todo[inline]{priklad?}

Autori preto zavádzajú key scoring function. If a new key $K$ is going to be inferred, it has assigned an initial score 0. Each inferred positive statement about $K$ increases the score of $K$ by the weight of the statement. Respectively, each negative statement about $K$ decreases the score by the respective weight. The resulting score therefore summarizes the weights of all inferred statements about $K$. A positive score means that $K$ is probably satisfied while negative means that $K$ is probably not satisfied. The higher the absolute value of the score is the higher the probability is.

@@

\section{Conclusion}
The output of our method is a list of scored keys and for each key a list foreign keys referencing the key. The score of a key can be negative or positive. A negative score means that the key is not specified by the XML documents while positive means that it is satisfied. The absolute value of the score means how sure the method are about it.

>>>>

\chapter{Analysis of Recent Approaches}
\todo[inline]{Translate.}
Many approaches of XML Schema inference have been proposed. Most of them process XML data as an input for inference. Some of these methods are mentioned in this chapter.

Besides approaches exploiting XML data, approaches that utilize operations over XML data may be developed and they are of the most interest in terms of this work. According to my best knowledge at the time of writing there is just one approach of this category. It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury?}.

\section{Discovering XML Keys and Foreign Keys in Queries}
This methods tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries. Just these queries are utilized in inference, no XML data are used. The output of this method is a set of keys and foreign keys that can be captured using XML Schema \textbf{key}, \textbf{keyref} \todo{aj unique?} constructs.

\subsection{Keys and foreign keys}
Authors introduce formalism that formalizes XML Schema keys.

\begin{define}[Key]
A key is a construct $$(C, P, \{L\})$$ where $C$, $P$, $L$ are XPath paths without predicates thet use only \texttt{child} and \texttt{descendant} axes.
$C$ can be omitted. $(P, \{L\})$ is then equivalent to $(/, P, \{L\})$ and is called \emph{global key}.Otherwise it is called \emph{local key}.
\end{define}

@@

\begin{define}[Foreign key]
A foreign key is a construct $$(C, (P_1, \{L_1\}) \Rightarrow (P_2, \{L_2\}))$$ where $(C, P_2, \{L_2\})$ is a key, $P_1$, $L_1$ are XPath paths without predicates thet use only \texttt{child} and \texttt{descendant} axes.
$C$ can be omitted as in the case of keys.
\end{define}

@@

\subsection{Key and foreign key discovery}
\subsubsection{Joins in queries}
To discover keys and foreign keys, method utilizes element/element joins.

@@

Assuming that each join is done via a key/foreign key pair, it can be inferred from $Q$ that $L_1$ is a key for elements in $S_1$ or $L_2$ is a key for elements in $S_2$ and the other is a foreign key referencing the key.

Authors introduces two observations:
\begin{enumerate}
\renewcommand{\theenumi}{(O\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\item Assume that an element $e_1$ from $S_1$ can be joined with more elements from $S_2$. It means that there can be more different elements in $S_2$ having their value of $L_2$ equal to $e_1/L_1$. In other words, there can be more different elements in $S_2$ having the same value of $L_2$. Therefore, we infer that $L_2$ can not be a key of the elements in $S_2$. Moreover, because we suppose that one of $L_1$ and $L_2$ is a key and $L_2$ can not be a key, we infer that $L_1$ is a key of the elements in $S_1$. Consequently, we infer that $L_2$ is a foreign key referring $L_1$.
\item Vice versa, assume that $e_1$ can be joined with maximally one element from $S_2$. It means that there is maximally one element in $S_2$ having its value of $L_2$ equal to $e_1/L_1$. In that case we suppose that each element in $S_2$ has a unique value of $L_2$. Therefore, we infer that $L_2$ is a key of the elements in $S_2$ and consequently that $L_1$ is a foreign key referring $L_2$. We cannot infer whether $L_1$ is a key or not.
\end{enumerate}

\subsubsection{Join patterns}
For a certain join, the decision for one of the cases (O1) and (O2) is made by the form of the join. The query is searched for so-called \emph{join patters}. These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and look like the following (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):
\todo[inline]{spravne kreslit indexy pri pismenkach}
\begin{verbatim}
01  for e1 in P1               for $e1 in P1
02  return                     return
03    for $e2 in                 let $e2 :=
        P2[L2 = $e1/L1]            P2[L2 = $e1/L1]
04    return CR                  return CR
\end{verbatim}

@@

\section{Scoring function}
It can happen that some of the assumptions the method is based on are not met. Assuming a set of queries, a key $K$ may be inferred from some query and processing of another query may result to denial of $K$ as a key. \todo[inline]{priklad?}

Therefore, authors introduce a scoring function. If a new key $K$ is going to be inferred, it has assigned an initial score 0. Each inferred positive statement about $K$ increases the score of $K$ by the weight of the statement. Respectively, each negative statement about $K$ decreases the score by the respective weight. The resulting score therefore summarizes the weights of all inferred statements about $K$. A positive score means that $K$ is probably satisfied while negative means that $K$ is probably not satisfied. The higher the absolute value of the score is the higher the probability is.

@@

\section{Conclusion}
The output of our method is a list of scored keys and for each key a list foreign keys referencing the key. The score of a key can be negative or positive. A negative score means that the key is not specified by the XML documents while positive means that it is satisfied. The absolute value of the score means how sure the method is about it.
